exponential_BT

### 2025-04-03 inst/exponential_BT.Rmd
more work

```{r makeII,cache=TRUE}
r_try <- 1:9
x_try <- seq(from = 0.1, to = 1, by = 0.1)
jj <- as.matrix(expand.grid(r_try, x_try))
II <- t(apply(jj,1,function(v){getprob(9,v[1],v[2],N=1e4)}))
head(II)
```

@@

support_func <- function(o,v){
    v <- v[-(1:2)]
    v <- v/sum(v)
    dmultinom(o[-(1:2)], prob=v, log=TRUE)
}
support <- rep(0,nrow(II))
for(i in seq_len(nrow(II))){
    support[i] <- support_func(o,II[i,])
}
```

```{r}
support <- support-max(support)
support <- pmax(support,-10)
support  <- matrix(support,length(r_try),length(x_try))
contour(r_try, x_try, support, xlab="r", ylab="x", levels = -(1:5)*2)
abline(v = r_true)
abline(h = x_true)
abline(v = 1:9, col = 'gray', lty = 3, lwd = 0.3)
filled.contour(r_try, x_try, support, xlab = "r", ylab = "x")
```

@@

We need a likelihood function for the observation (9) of $\alpha,x$.
Let's try again.

```{r ninetwelve, cache=TRUE}
n <- 12
a_try <- seq(from=0.01,to=0.99,len=10)
p_try <- ceiling(n * a_try)
x_try <- seq(from = 0.1, to = 1, by = 0.1)
jj <- as.matrix(expand.grid(p_try, x_try))
II <- t(apply(jj,1,function(v){getprob(n, v[1], v[2], N=1e4)}))
head(II)
tail(II)
S912 <- matrix(II[,11],length(a_try),length(x_try))
```

```{r show912}
S912 <- S912/max(S912)
contour(p_try,a_try,log(S912))
```

# Variable class sizes

Let's try this on the apocryphal education dataset

```{r}
M <- matrix(c(
    9, 7, 2, 3, 2, 8, 3, 4, 9,
    12,17,23,09,13,14,13,12,15
),nrow=2,byrow=TRUE)
M
```

>>>>

```{r makeII,cache=TRUE}
r_try <- 1:9
x_try <- seq(from = 0.1, to = 1, by = 0.1)
jj <- as.matrix(expand.grid(r_try, x_try))
II <- t(apply(jj, 1, function(v){getprob(9, v[1], v[2], N=1e4)}))
head(II)
```

@@

support_func <- function(o, v){
    v <- v[-(1:2)]
    v <- v/sum(v)
    dmultinom(o[-(1:2)], prob=v, log=TRUE)
}
support <- rep(0,nrow(II))
for(i in seq_len(nrow(II))){
    support[i] <- support_func(o,II[i,])
}
```

```{r}
support <- support-max(support)
support <- pmax(support, -10)
support  <- matrix(support,length(r_try),length(x_try))
contour(r_try, x_try, support, xlab="r", ylab="x", levels = -(1:5)*2)
abline(v = r_true)
abline(h = x_true)
abline(v = 1:9, col = 'gray', lty = 3, lwd = 0.3)
filled.contour(r_try, x_try, support, xlab = "r", ylab = "x")
```

@@

```{r ninetwelve}
a_try <- seq(from=0.01,to=0.99,len=10)
x_try <- seq(from = 0.1, to = 1, by = 0.1)

single_like_fun <- function(n, a, N=1e4){
    p_try <- ceiling(n * a_try)
    jj <- as.matrix(expand.grid(p_try, x_try))
    II <- t(apply(jj,1,function(v){getprob(n, v[1], v[2], N)}))
    matrix(II[, ceiling(a*n) + 2], length(a_try), length(x_try))
}
```

```{r makesupp912, cache=TRUE}
allsupp <- list(
    log(single_like_fun(12, a =  8/12)),  # pure
    log(single_like_fun(17, a = 12/17)),  # pure
    log(single_like_fun(23, a = 18/23)),  # pure
    log(single_like_fun(14, a =  5/14)),  # applied
    log(single_like_fun(13, a =  3/13)),  # applied
    log(single_like_fun(11, a =  2/11)),  # applied
    log(single_like_fun(13, a =  3/13))   # applied
)
```

```{r showallsupp}
S <- Reduce("+",allsupp)
S <- pmax(S-max(S),-10)
contour(a_try, x_try, S, levels = -1*(1:7))
jj <- which(S == max(S), arr.ind = TRUE)
points(a_try[jj[1]], x_try[jj[2]], pch=16, col = 'red')
```

```{r showbothsupp}
S_pure    <- Reduce("+", allsupp[1:3])
S_applied <- Reduce("+", allsupp[4:7])
S_all     <- Reduce("+", allsupp[1:7])
max(S_all)
max(S_pure)
max(S_applied)
max(S_pure) + max(S_applied)
max(S_pure) + max(S_applied) - max(S_all)
pchisq(max(S_pure) + max(S_applied) - max(S_all),df=2,lower.tail=FALSE)
```

# Parkrun

```{r parkrunlikefun1, cache=TRUE}

# In this chunk, each execution of single_like_fun(...,N=100) takes ~1
# minute

 print(system.time(parkrun01 <- single_like_fun(259, 177/259, N=1e3)))
 print(system.time(parkrun02 <- single_like_fun(305, 222/305, N=1e3)))
 print(system.time(parkrun03 <- single_like_fun(297, 206/297, N=1e3)))
#print(system.time(parkrun04 <- single_like_fun(241, 142/241, N=1e1)))
#print(system.time(parkrun05 <- single_like_fun(179, 118/179, N=1e1)))
#print(system.time(parkrun06 <- single_like_fun(338, 224/338, N=1e1)))
#print(system.time(parkrun07 <- single_like_fun(203, 128/203, N=1e1)))
```

```{r}
parkrun01
parkrun02
parkrun03
#parkrun04
#parkrun05
#parkrun06
#parkrun07
sum(parkrun01)
sum(parkrun02)
sum(parkrun03)
parkrun01 * parkrun02 *parkrun03
sum(parkrun01 * parkrun02 *parkrun03)
```
